### 2017-10-09 proposal.Rmd
proposal.Rmd: initial commit

---
title: "Proposal"
author: "Mark H. White II | markhwhiteii@gmail.com"
output: pdf_document
---