### 2017-11-30 docs/presentation_handout.Rmd
initial commit

---
output: pdf_document
---

```{r warning = FALSE, message = FALSE, results = FALSE, echo = FALSE}
library(tidyverse)
library(ggridges)
library(knitr)
library(rstan)
dat <- read_csv("../data/results_df.csv") %>% 
  bind_rows(read_csv("../data/results_df_101-150.csv")) %>% 
  bind_rows(read_csv("../data/results_df_151-200.csv")) %>% 
  bind_rows(read_csv("../data/results_df_201-300.csv")) %>% 
  bind_rows(read_csv("../data/results_df_301-350.csv")) %>% 
  bind_rows(read_csv("../data/results_df_351-450.csv")) %>% 
  bind_rows(read_csv("../data/results_df_451-500.csv")) %>% 
  separate(v1, c("algorithm", "sampling"), "_", FALSE) %>% 
  mutate(
    v1 = as.factor(v1),
    algorithm = factor(
      algorithm, levels = c("c50", "adaboost", "randomforest", "xgboost")
      ),
    sampling = as.factor(sampling),
    predict_vars = linear_vars + 5,
    iter = as.factor(iter),
    spec = tn / (tn + fp),
    auc_roc = (1 + rec - (1 - spec)) / 2
  )
names(dat)[1] <- "model"
levels(dat$model) <- c(
  "AdaBoost, None", "AdaBoost, Over", "AdaBoost, SMOTE", "AdaBoost, Under",
  "C5.0, None", "C5.0, Over", "C5.0, SMOTE", "C5.0, Under",
  "Random Forest, None", "Random Forest, Over", 
  "Random Forest, SMOTE", "Random Forest, Under",
  "XGBoost, None", "XGBoost, Over", "XGBoost, SMOTE", "XGBoost, Under")
dat <- dat[, !names(dat) %in% c("cor_vars", "linear_vars")]
```

# Class Imbalance
- Minority class much smaller than majority  
- Class of interest ("positive class"), e.g., is only 5% of data  
- Problem: 95% accurate guessing all negative; not helpful algorithm

## Sampling Techniques

### Over
- Randomly replicate minority class samples until training data is balanced

### Under
- Randomly throw out majority class samples until training data is balanced

### SMOTE
- Create synthetic minority cases, based on $k$-nearest neighbors

## Ensembling Algorithms

### Bagging
- In this study, using random forest  
- Bootstrap the training data, select a subset of predictors, train decision tree on this data  
- Do this a given number of times  
- Prediction is majority vote of all of these decision trees

### Boosting
- In this study, using AdaBoost and XGBoost  
- Both train trees *serially*  
- Learn from the mistakes of past trees by updating weights based on mistakes or updating based on the residuals using gradient descent

# Data Generation
- Two multivariate normal predictors ($A$ and $B$) are generated. $A$ and $B$ are correlated at $r = .65$. These two variables contributed to the log-odds by $4A + 4B + 2AB$  
- Another variable, $J \sim U(-1, 1)$, was generated. This variable further added to the log-odds by $J^3 + 2 \times \exp(-6 \times (J - 0.3)^2)$  
- Two more variables, $K \sim U(0, 1)$ and $L \sim U(0, 1)$, were generated and contributed to the log-odds by $2 \times \sin(K \times L)$  
- For each data set, a number $X$ was selected, where $X \sim N(50, 7)$. Another number, $Y$, was selected, where $Y \sim N(.15, .033)$. $Z = X - (X \times Y)$ variables were generated from a $N(0, 1)$ distribution. Each of these $Z$ variables further added to the log-odds in a simple additive fashion, where coefficients were (a) of alternating signs and (b) evenly spaced from $2.50$ to $0.25$  
- $Y \over 2$ variables were generated from a $N(0, 1)$ distribution and did not contribute to the log-odds  
- The log-odds for each case were converted to probabilities. For each data set, a positive (i.e., minority) class proportion, $M$, was sampled from $N(.03, .007)$. Probabilities were sorted from lowest to highest. The difference between the probability for the $1 - M$th highest probability and $M$ was calculated, and this constant was added to the probability for each case  
- Lastly, the number of cases in each data set were randomly drawn from a distribution $N(40000, 5000)$.  500 data sets were generated, and sixteen combinations of sampling techniques and algorithms were fit to each of these data sets

# Performance Assessment

## Precision
- $TP \over TP + FP$

## Recall
- $TP \over TP + FN$

## F1
- $F_1 = 2 \times {\text{precision} \times \text{recall} \over \text{precision} + \text{recall}}$

## AUC(ROC)
- $\text{AUC(ROC)} = {1 + \text{recall} - \text{false positive rate} \over 2}$

```{r echo = FALSE, out.width = "65%", fig.align = "center"}
roc <- data.frame(x = c(0, 1, 0, .4, 1),
                  y = c(0, 1, 0, .6, 1),
                  z = c(1, 1, 0, 0, 0))
ggplot(roc, aes(x = x, y = y, linetype = factor(z))) +
  geom_line() +
  geom_point(aes(x = .4, y = .6)) +
  geom_point(aes(x = .5, y = .5)) +
  theme_minimal() +
  theme(legend.position = "none") +
  labs(x = "False Positive Rate", y = "Recall") +
  annotate("text", x = .35, y = .70, label = "AUC = .60 \n (.40, .60)")
```

# Results

## Making Enough Positive Predictions
```{r echo = FALSE}
prop_zeros <- dat %>% 
  mutate(total_pos = tp + fp) %>% 
  group_by(model) %>% 
  summarise_at("total_pos", funs(mean(. == 0))) %>% 
  arrange(desc(total_pos))
kable(prop_zeros, col.names = c("Model", "Proportion P = 0"))
```

```{r echo = FALSE}
prop_f1_nan <- dat %>% 
  mutate(f1_nan = is.nan(f1)) %>% 
  group_by(model) %>% 
  summarise_at("f1_nan", funs(mean(. == TRUE))) %>% 
  arrange(desc(f1_nan))
kable(prop_f1_nan, col.names = c("Model", "Proportion F1 is N/A"))
```

## Comparing Mean Performance
```{r echo = FALSE}
good_models <- prop_f1_nan %>% 
  filter(f1_nan == 0) %>% 
  pull(model) %>% 
  as.character()
mean_results <- dat %>% 
  filter(model %in% good_models) %>% 
  select(model, prec, rec, f1, auc_roc) %>% 
  group_by(model) %>% 
  summarise_if(is.numeric, mean) %>% 
  arrange(desc(auc_roc))
kable(mean_results, digits = 3, col.names = 
        c("Model", "Precision", "Recall", "F1", "AUC(ROC)"))
```

```{r echo = FALSE, warning = FALSE, message = FALSE}
good_dat <- dat[dat$model %in% good_models, ]
ranked_names <- names(sort(tapply(
  good_dat$auc_roc, droplevels(good_dat$model), mean, na.rm = TRUE),
  decreasing = TRUE
))
good_dat$model <- factor(good_dat$model, levels = c(ranked_names))
dat_long_ridge <- good_dat %>% 
  transmute(
    model = model, 
    `AUC(ROC)` = auc_roc, 
    F1 = f1, 
    Precision = prec, 
    Recall = rec) %>% 
  gather(metric, value, -model)
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", 
               "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
ggplot(dat_long_ridge, aes(y = model, x = value, fill = model)) +
  geom_density_ridges(alpha = .7) +
  facet_wrap( ~ metric, scales = "free_x") +
  theme_minimal() +
  labs(x = NULL, y = "Model") +
  theme(text = element_text(size = 14), legend.position = "none") +
  scale_fill_manual(values = cbPalette)
```

```{r warning = FALSE, message = FALSE, results = FALSE, echo = FALSE}
stan_code <- "
data {
  int n; // observations
  int k; // number of models
  real gm; // grand mean
  real psigmu; // prior for mu_sigma
  int x[n]; // group indicator
  vector[n] y; // outcome
}
parameters {
  real alpha[k]; // model means
  real<lower=0> sigma; // error
}
transformed parameters {
  real rf_xg_diff; // compare rf and xg
  real rf_ada_diff; // compare rf and ada
  real xg_ada_diff; // compare xg and ada
  rf_xg_diff = alpha[1] - alpha[2]; // compute diff
  rf_ada_diff = alpha[1] - alpha[3]; // compute diff
  xg_ada_diff = alpha[2] - alpha[3]; // compute diff
}
model {
  vector[n] mu; // expected value
  for (i in 1:n) { // for all people
    mu[i] = alpha[x[i]]; //their mu is their group's mean
  }
  y ~ normal(mu, sigma); // likelihood
  alpha ~ normal(gm, psigmu); // prior, defined from data
  sigma ~ cauchy(0, 1); // uninformative prior on sigma
}
"
dat_long <- good_dat %>% 
  mutate(model = as.numeric(model)) %>% 
  select(model, auc_roc, f1, prec, rec)
set.seed(1839)
sdat_aucroc <- list(n = nrow(dat_long), k = length(unique(dat_long$model)), 
                    gm = mean(dat_long$auc_roc), psigmu = .10, x = dat_long$model,
                    y = dat_long$auc_roc)
sdat_f1 <- list(n = nrow(dat_long), k = length(unique(dat_long$model)), 
                    gm = mean(dat_long$f1), psigmu = .005, x = dat_long$model,
                    y = dat_long$f1)
sdat_prec <- list(n = nrow(dat_long), k = length(unique(dat_long$model)), 
                    gm = mean(dat_long$prec), psigmu = .005, x = dat_long$model,
                    y = dat_long$prec)
sdat_rec <- list(n = nrow(dat_long), k = length(unique(dat_long$model)), 
                    gm = mean(dat_long$rec), psigmu = .10, x = dat_long$model,
                    y = dat_long$rec)
mdiff_aucroc <- stan(model_code = stan_code, data = sdat_aucroc,
                     control = list(max_treedepth = 15))
mdiff_f1 <- stan(model_code = stan_code, data = sdat_f1,
                 control = list(max_treedepth = 15))
mdiff_prec <- stan(model_code = stan_code, data = sdat_prec,
                   control = list(max_treedepth = 15))
mdiff_rec <- stan(model_code = stan_code, data = sdat_rec,
                  control = list(max_treedepth = 15))
mean_diffs <- round(summary(mdiff_aucroc)$summary[10:12, c(1, 4, 8)], 3) %>% 
  rbind(round(summary(mdiff_f1)$summary[10:12, c(1, 4, 8)], 3)) %>% 
  rbind(round(summary(mdiff_rec)$summary[10:12, c(1, 4, 8)], 3)) %>% 
  rbind(round(summary(mdiff_prec)$summary[10:12, c(1, 4, 8)], 3)) %>% 
  as.data.frame() %>% 
  rownames_to_column("Pairwise Comparison") %>% 
  mutate(`Pairwise Comparison` = rep(c(
        "Random Forest - XGBoost", 
        "Random Forest - AdaBoost",
        "XGBoost - AdaBoost"), 4),
         Outcome = c("AUC(ROC)", "", "", "F1", "", "",
                     "Recall", "", "", "Precision", "", ""))
mean_diffs <- mean_diffs[, c(5, 1, 2:4)]
colnames(mean_diffs)[3] <- "Difference"
```
``` {r echo = FALSE}
kable(mean_diffs)
```

## Performance With Data Characteristics
```{r echo = FALSE}
auc_cor_dat <- dat[dat$model %in% good_models, ] %>% 
  transmute(Model = model, `AUC(ROC)` = auc_roc, N = n,  
            `Minority Size` = minority_size, Predictors = predict_vars,
            `Noise Variables` = noise_vars) %>% 
  gather("key", "value", 3:6)
ggplot(auc_cor_dat, 
       aes(x = value, y = `AUC(ROC)`, group = Model, color = Model)) +
  geom_point(alpha = .4) +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ key, scales = "free_x") +
  labs(x = NULL) +
  theme_minimal() +
  scale_colour_manual(values = cbPalette)
```